---
title: "Reproducible Research: Peer Assessment 1"
author: "Hong Chen"
Date: "March 6, 2020"
output: html_document
keep_md: true
---
library(dplyr)
library(ggplot2)


## Loading and preprocessing the data

```{r dataprep, echo=TRUE, include=TRUE}
  data<- read.csv(unzip("activity.zip"))
  names(data)
  #head(data)
  
```


## What is mean total number of steps taken per day?
```{r statperday, echo=TRUE, include=TRUE}
  step_sum <- aggregate(steps~date, data, sum)
  head(step_sum)
  hist(step_sum$steps,main = "Histogram of Daily Steps", xlab="Steps", )
  step_mean <- mean(step_sum$steps)
  step_median <- median(step_sum$steps)
  print(paste("The mean is: ", step_mean))
  print(paste("The median is: ", step_median))
```

## What is the average daily activity pattern?
```{r statperinteval, echo=TRUE, include=TRUE}
  step_interval <- aggregate(steps~interval, data, mean)
  head(step_interval)
  plot(step_interval$steps ~ step_interval$interval,
     type="l", xlab = "5 Minute Intervals", ylab = "Average Number of Steps",
     main = "Steps By Time Interval")
  Max_steps_interval <- step_interval$interval[which.max(step_interval$steps)]
  print(paste("Interval containing the most steps on average: ",Max_steps_interval))
  
```
  

## Imputing missing values
```{r missingvalue, echo=TRUE, include=TRUE}
  NA_sum <- sum(is.na(data$steps))
  print(paste("The total number of rows with NA is: ",NA_sum))
  new_data <- data  
  for (i in 1:nrow(data)){
        if(is.na(data$steps[i])){
                new_data$steps[i]<- step_interval$steps[new_data$interval[i] == step_interval$interval]
        }
  }
  head(new_data)
  new_step_sum <- aggregate(steps~date, new_data, sum)
  hist(new_step_sum$steps,main = "Histogram of Daily Steps", xlab="Steps", )
  new_step_mean <- mean(new_step_sum$steps)
  new_step_median <- median(new_step_sum$steps)
  print(paste("New mean is: ", new_step_mean))
  print(paste("New median is: ", new_step_median))

  

## Are there differences in activity patterns between weekdays and weekends?
